Preparación de datos

### 2022-06-15 production/prod_prep.Rmd
update ola cateogorica y % migrant

#
sjmisc::frq(elsoc_long_2016_2021$in_simpat)
elsoc_long_2016_2021$in_simpat <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_simpat, 
                      label = "Degree of sympathy for [PER/HAI/VEN] living in Chile") 
elsoc_long_2016_2021$in_simpat <-
sjlabelled::set_labels(x = elsoc_long_2016_2021$in_simpat,
                       labels = c("Very little or not at all", "A little",
                                  "Somewhat","Quite a lot","A lot"))  
sjmisc::frq(elsoc_long_2016_2021$in_simpat)

#
sjmisc::frq(elsoc_long_2016_2021$in_simili)
elsoc_long_2016_2021$in_simili <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_simili, 
                      label = "Degree of similarity between Chileans and [PER/HAI/VEN].") 
elsoc_long_2016_2021$in_simili <-
sjlabelled::set_labels(x = elsoc_long_2016_2021$in_simili,
                       labels = c("Not similar", "Slightly similar", "Somewhat similar",
                                  "Quite similar", "Very similar"))
sjmisc::frq(elsoc_long_2016_2021$in_simili)

@@

sjmisc::frq(elsoc_long_2016_2021$in_valfam)
elsoc_long_2016_2021$in_valfam <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_valfam, 
                      label = "Degree of agreement: My family values my friends [PER/HAI/VEN].") 
sjmisc::frq(elsoc_long_2016_2021$in_valfam)

sjmisc::frq(elsoc_long_2016_2021$in_valami)
elsoc_long_2016_2021$in_valami <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_valami, 
                      label = "Degree of agreement: My friends value my friends [PER/HAI/VEN].") 
sjmisc::frq(elsoc_long_2016_2021$in_valami)

sjmisc::frq(elsoc_long_2016_2021$in_amsimb)
elsoc_long_2016_2021$in_amsimb <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_amsimb, 
                      label = "Degree of agreement: Chile loses its identity with the arrival of [PER/HAI/VEN].") 
sjmisc::frq(elsoc_long_2016_2021$in_amsimb)

sjmisc::frq(elsoc_long_2016_2021$in_amreal)
elsoc_long_2016_2021$in_amreal <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_amreal, 
                      label = "Degree of agreement: With the arrival of [PER/HAI/VEN] unemployment rises.") 
sjmisc::frq(elsoc_long_2016_2021$in_amreal)

sjmisc::frq(elsoc_long_2016_2021$in_acpier)
elsoc_long_2016_2021$in_acpier <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_acpier, 
                      label = "Degree of agreement: [PER/HAI/VEN] keep their customs") 
sjmisc::frq(elsoc_long_2016_2021$in_acpier)

sjmisc::frq(elsoc_long_2016_2021$in_acadop)
elsoc_long_2016_2021$in_acadop <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_acadop, 
                      label = "Degree of agreement: [PER/HAI/VEN] adopt Chilean customs") 
sjmisc::frq(elsoc_long_2016_2021$in_acadop)

sjmisc::frq(elsoc_long_2016_2021$in_amigcl)
elsoc_long_2016_2021$in_amigcl <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_amigcl, 
                      label = "Degree of agreement: [PER/HAI/VEN] have Chilean friends") 
sjmisc::frq(elsoc_long_2016_2021$in_amigcl)

@@

df_study1 <- 
elsoc_long_2016_2021 %>% 
  filter(muestra == 1) %>% #keep original sample (vs refresh)
  select(idencuesta,ola,region,region_cod,comuna,comunacod=comuna_cod,
         muestra,
         cuestion_mig,
         starts_with("in_"),
         educ,educyear,
         # ing_pc,
         # quintil,
         quintil1,
         ess,
         # ess_inm,
         nation_t1,
         # know_inm,
         # know_inm_bi,
         # frien_inm_bi_t4,
         # frien_inm_bi_t1,
         know_inm_t1t4,
         # frien_inm,
         # frien_inm_bi,
         frien_inm_t1t4,
         pos_id,
         edad)

@@

# reshape from long to wide
pacman::p_load(datasets,data.table)
df_study1_long <- data.table::melt.data.table(data.table::setDT(df_study1_wide),
              id.vars = c("idencuesta","muestra"),
              variable.name = c("ola"),
              measure = patterns("^comuna_","^comunacod_","^cuestion_mig_",
                                 "^in_simpat_","^in_simili_","^in_ansied_","^in_valfam_",
                                 "^in_valami_","^in_amsimb_","^in_amreal_","^in_acpier_",
                                 "^in_acadop_","^in_amigcl_",
                                 "^educ_","^educyear_","^quintil1_","^ess_","^nation_t1_",
                                 "^know_inm_t1t4_","^frien_inm_t1t4_","^pos_id_",
                                 "^edad_"
                                 ),
              value.name = c("comuna","comunacod","cuestion_mig",
                             "in_simpat","in_simili","in_ansied","in_valfam",
                             "in_valami","in_amsimb","in_amreal",
                             "in_acpier","in_acadop","in_amigcl",
                             "educ","educyear","quintil1","ess","nation_t1",
                             "know_inm_t1t4","frien_inm_t1t4","pos_id",
                             "edad"),
              na.rm = F,value.factor = TRUE
              )

>>>>

#
sjmisc::frq(elsoc_long_2016_2021$in_simpat)
elsoc_long_2016_2021$in_simpat <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_simpat, 
                      label = "Sympathy for [PER/HAI/VEN] living in Chile") 
elsoc_long_2016_2021$in_simpat <-
sjlabelled::set_labels(x = elsoc_long_2016_2021$in_simpat,
                       labels = c("Very little or not at all", "A little",
                                  "Somewhat","Quite a lot","A lot"))  
sjmisc::frq(elsoc_long_2016_2021$in_simpat)

#
sjmisc::frq(elsoc_long_2016_2021$in_simili)
elsoc_long_2016_2021$in_simili <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_simili, 
                      label = "Similarity between Chileans and [PER/HAI/VEN]") 
elsoc_long_2016_2021$in_simili <-
sjlabelled::set_labels(x = elsoc_long_2016_2021$in_simili,
                       labels = c("Not similar", "Slightly similar", "Somewhat similar",
                                  "Quite similar", "Very similar"))
sjmisc::frq(elsoc_long_2016_2021$in_simili)

@@

sjmisc::frq(elsoc_long_2016_2021$in_valfam)
elsoc_long_2016_2021$in_valfam <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_valfam, 
                      label = "Family value of immigrant friends [PER/HAI/VEN]") 
sjmisc::frq(elsoc_long_2016_2021$in_valfam)

sjmisc::frq(elsoc_long_2016_2021$in_valami)
elsoc_long_2016_2021$in_valami <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_valami, 
                      label = "Friends value of immigrant friends [PER/HAI/VEN]") 
sjmisc::frq(elsoc_long_2016_2021$in_valami)

sjmisc::frq(elsoc_long_2016_2021$in_amsimb)
elsoc_long_2016_2021$in_amsimb <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_amsimb, 
                      label = "Loose of identity because of immigrants") 
sjmisc::frq(elsoc_long_2016_2021$in_amsimb)

sjmisc::frq(elsoc_long_2016_2021$in_amreal)
elsoc_long_2016_2021$in_amreal <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_amreal, 
                      label = "Unemployment increases") 
sjmisc::frq(elsoc_long_2016_2021$in_amreal)

sjmisc::frq(elsoc_long_2016_2021$in_acpier)
elsoc_long_2016_2021$in_acpier <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_acpier, 
                      label = "Immigrant keeping their customs") 
sjmisc::frq(elsoc_long_2016_2021$in_acpier)

sjmisc::frq(elsoc_long_2016_2021$in_acadop)
elsoc_long_2016_2021$in_acadop <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_acadop, 
                      label = "Adoption of Chilean customs") 
sjmisc::frq(elsoc_long_2016_2021$in_acadop)

sjmisc::frq(elsoc_long_2016_2021$in_amigcl)
elsoc_long_2016_2021$in_amigcl <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$in_amigcl, 
                      label = "Immigrants [PER/HAI/VEN] have Chilean friends") 
sjmisc::frq(elsoc_long_2016_2021$in_amigcl)

# Average variables___________________________________________________________

elsoc_long_2016_2021 <- elsoc_long_2016_2021 %>% 
  mutate(conviviality=as.numeric((in_ansied+in_simpat+in_valfam+in_valami+in_amigcl)/5),
         identity= as.numeric((in_simili+in_amsimb+in_acpier+in_acadop)/4),
         threat = as.numeric(in_amreal))

sjmisc::frq(elsoc_long_2016_2021$conviviality)
elsoc_long_2016_2021$conviviality <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$conviviality, 
                      label = "Conviviality")
sjmisc::frq(elsoc_long_2016_2021$conviviality)

sjmisc::frq(elsoc_long_2016_2021$identity)
elsoc_long_2016_2021$identity <- 
sjlabelled::set_label(x = elsoc_long_2016_2021$identity, 
                      label = "Identity")
sjmisc::frq(elsoc_long_2016_2021$identity)

@@

df_study1 <- 
elsoc_long_2016_2021 %>% 
  filter(muestra == 1) %>% #keep original sample (vs refresh)
  select(idencuesta,ola,region,region_cod,comuna,comunacod=comuna_cod,
         muestra,
         cuestion_mig,
         starts_with("in_"),
         conviviality,
         identity,
         educ,educyear,
         # ing_pc,
         # quintil,
         quintil1,
         ess,
         # ess_inm,
         nation_t1,
         # know_inm,
         # know_inm_bi,
         # frien_inm_bi_t4,
         # frien_inm_bi_t1,
         know_inm_t1t4,
         # frien_inm,
         # frien_inm_bi,
         frien_inm_t1t4,
         pos_id,
         edad)

@@

# reshape from long to wide
pacman::p_load(datasets,data.table)
df_study1_long <- data.table::melt.data.table(data.table::setDT(df_study1_wide),
              id.vars = c("idencuesta","muestra"),
              variable.name = c("ola"),
              measure = patterns("^comuna_","^comunacod_","^cuestion_mig_",
                                 "^in_simpat_","^in_simili_","^in_ansied_","^in_valfam_",
                                 "^in_valami_","^in_amsimb_","^in_amreal_","^in_acpier_",
                                 "^in_acadop_","^in_amigcl_",
                                 "^conviviality","^identity",
                                 "^educ_","^educyear_","^quintil1_","^ess_","^nation_t1_",
                                 "^know_inm_t1t4_","^frien_inm_t1t4_","^pos_id_",
                                 "^edad_"
                                 ),
              value.name = c("comuna","comunacod","cuestion_mig",
                             "in_simpat","in_simili","in_ansied","in_valfam",
                             "in_valami","in_amsimb","in_amreal",
                             "in_acpier","in_acadop","in_amigcl",
                             "conviviality","identity",
                             "educ","educyear","quintil1","ess","nation_t1",
                             "know_inm_t1t4","frien_inm_t1t4","pos_id",
                             "edad"),
              na.rm = F,value.factor = TRUE
              )
